---
title: "univariate_RV(c)+RV(d)"
author: "CID: 01714151"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Define functions to distinct jump and diffusive parts on thresholds
```{r}
summary3 <- function(dat_flat, time_func, jump_threshold){
  dat_summary <- dat_flat %>%
    # Add a column which represents the date/month of the dataset
    mutate(date=time_func(datetime)) %>%
    # Change the index and group by date
    index_by(date) %>%
    group_by(date) %>%
    # get the average price, RVol, P(RVol), N(RVol)
    summarize(currency = first(currency),
              avg_price = mean(price),
              rvol = sqrt(sum(log_return^2)),
              #BV_threshold = sqrt(sum(bpv)*mu1^(-2)),
              rvol_c = sqrt(sum(log_return[abs(log_return) <= jump_threshold]^2)),
              rvol_d = sqrt(sum(log_return[abs(log_return) > jump_threshold]^2)),
              P_rvol_c = sqrt(sum(log_return[(log_return>0) & (abs(log_return) <= jump_threshold)]^2)),
              P_rvol_d = sqrt(sum(log_return[(log_return>0) & (abs(log_return) > jump_threshold)]^2)),
              N_rvol_c = sqrt(sum(log_return[(log_return<0) & (abs(log_return) <= jump_threshold)]^2)),
              N_rvol_d = sqrt(sum(log_return[(log_return<0) & (abs(log_return) > jump_threshold)]^2)))
  return(dat_summary)
}

#const_threshold<- function(const){return(const)}
#BV_threshold<- function(alpha){alpha*sqrt(EU_daily2$rbpv*mu1^(-2))*(1/288)^0.49}
```

## HAR
### Seperate and explore RV_c and RV_d
```{r}
# Obtain daily summary for FX pairs which include the test statistic and corresponding critival values
EU_daily3 <- summary3(EU_flat, lubridate::date, jump_threshold = 0.002)
#AU_daily3 <- summary3(AU_flat, lubridate::date, jump_threshold = 0.005)
length(EU_daily3$rvol_d[EU_daily3$rvol_d != 0])

acf(EU_daily3$rvol_c, lag.max=30, main="EURUSD Daily Realized Vol (continous), 2022 May - 2023 Apr")
acf(EU_daily3$rvol_d, lag.max=30, main="EURUSD Daily Realized Vol (discontinous), 2022 May - 2023 Apr")
```
### HAR_RV_c in-sample forecasting for $\sigma^2$
```{r}
EU_HAR_in_model_c <- comp_har_pred(23, 256, EU_daily3$rvol_c)$model
EU_HAR_in_model_c$coefficients
base::summary(EU_HAR_in_model_c)$adj.r.squared
EU_HAR_inpred_c <- data.frame(date=EU_daily$date[24:257], pred = predict(EU_HAR_in_model_c), true = EU_daily3$rvol_c[24:257])

# Plot the HAR in-sample forecast
EU_HAR_inpred_c_plot <- ggplot(data=EU_HAR_inpred_c, aes(x=date)) +
  geom_line(aes(y=true, color='True RVol_c')) +
  geom_line(aes(y=pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('True RVol_c' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("RVol") +
  theme_classic()

EU_HAR_inpred_c_plot
```
### HAR_RV_c Out-of-sample forecasting for $\sigma^2$
```{r echo=FALSE}
# Obtain the one-day out-of-sample forecast and 0.95 CI of the rolling window
EU_HAR_outpred_c <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_har_pred(x-train_length, x, EU_daily3$rvol_c)$pred})), row.names = NULL)
EU_HAR_outpred_c$date = EU_daily3$date[(EU_roll_start+2) : (EU_roll_end+2)]
EU_HAR_outpred_c$true = EU_daily3$rvol_c[(EU_roll_start+2) : (EU_roll_end+2)]

mean((EU_HAR_inpred_c$pred - EU_HAR_inpred_c$true)^2)
mean((EU_HAR_outpred_c$true - EU_HAR_outpred_c$fit)^2)

# Plot the HAR out-of-sample forecast
EU_daily3_23 <- EU_daily3[year(EU_daily3$date)==2023,]
EU_HAR_outpred_plot_c <- ggplot() +
  geom_line(data=data.frame(x=EU_daily3_23$date, y=EU_daily3_23$rvol_c), 
            aes(x = x, y = y, color = "realized √BPV/µ1")) +
  geom_line(data=data.frame(x=EU_HAR_outpred_c$date, y=EU_HAR_outpred_c$fit), 
            aes(x = x, y = y, color = "HAR pred")) +
  geom_ribbon(data=data.frame(x=EU_HAR_outpred_c$date, lwr=EU_HAR_outpred_c$lwr, upr = EU_HAR_outpred_c$upr), aes(x = x, ymin = lwr, ymax=upr), fill = "lightblue", alpha=0.3) +
  scale_color_manual(name=NULL, values=c('realized √BPV/µ1' = 'darkgrey', "HAR pred" = "darkblue")) +
  theme_classic() +
  ggtitle("HAR Out-of-sample Forecast (with 95% CI) on diffusive parts of EURUSD from 2023 Mar - Apr") +
  xlab("date") + ylab("BPV")

EU_HAR_outpred_plot_c
```
### HAR, dHAR and naive in-sample prediction for J(ump)
```{r}
# Estimate the model HAR_J
EU_HAR_in_model_d <- comp_har_pred(23, 256, EU_daily3$rvol_d)$model
EU_HAR_in_model_d$coefficients
base::summary(EU_HAR_in_model_d)$adj.r.squared

# Estimate the model dHAR_J
EU_dHAR_in_model_d <- comp_dhar_pred(23, 256, EU_daily3$rvol_d)$model
EU_dHAR_in_model_d$coefficients
base::summary(EU_dHAR_in_model_d)$adj.r.squared

# Estimate the model Naive_J
EU_naive_in_d <- EU_daily3$rvol_d[23:256]
```
### HAR RV_c+RV_d in-sample forecast
```{r}
# HAR_BPV + HAR_J
EU_HAR_inpred_cd <- data.frame(date=EU_daily3$date[24:257], true=EU_daily3$rvol[24:257],
                               HAR_pred=sqrt(predict(EU_HAR_in_model_c)^2 +
                                               predict(EU_HAR_in_model_d)^2),
                               dHAR_pred=sqrt(predict(EU_HAR_in_model_c)^2 +
                                                predict(EU_dHAR_in_model_d)^2),
                               naive_pred=sqrt(predict(EU_HAR_in_model_c)^2 + EU_naive_in_d^2))

# Plot the HAR in-sample forecast
EU_HAR_inpred_plot_cd <- ggplot(data=EU_HAR_inpred_cd, aes(x=date)) +
  geom_line(aes(y=true, color='RVol')) +
  geom_line(aes(y=HAR_pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('RVol' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("BPV+J") +
  theme_classic()

EU_HAR_inpred_plot_cd

EU_dHAR_inpred_plot_cd <- ggplot(data=EU_HAR_inpred_cd, aes(x=date)) +
  geom_line(aes(y=true, color='RVol')) +
  geom_line(aes(y=dHAR_pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('RVol' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("BPV+J") +
  theme_classic()

EU_dHAR_inpred_plot_cd

EU_naive_inpred_plot_cd <- ggplot(data=EU_HAR_inpred_cd, aes(x=date)) +
  geom_line(aes(y=true, color='RVol')) +
  geom_line(aes(y=naive_pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('RVol' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("BPV+J") +
  theme_classic()

EU_naive_inpred_plot_cd
```
### HAR, dHAR and naive out-of-sample prediction for RV_d
```{r}
EU_HAR_outpred_d <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_har_pred(x-train_length, x, EU_daily3$rvol_d)$pred})), row.names = NULL)
EU_dHAR_outpred_d <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_dhar_pred(x-train_length, x, EU_daily3$rvol_d)$pred})), row.names = NULL)
EU_naive_outpred_d <- EU_daily3$rvol_d[(EU_roll_start+1): (EU_roll_end+1)]
```

### HAR RV_c+RV_d out-of-sample forecast
```{r}
# Obtain the one-day out-of-sample forecast of the rolling window
EU_HAR_outpred_cd <- data.frame(date=EU_daily3$date[(EU_roll_start+2) : (EU_roll_end+2)],
                                true=EU_daily3$rvol[(EU_roll_start+2) : (EU_roll_end+2)],
                                HAR_pred=sqrt(EU_HAR_outpred_c$fit^2+EU_HAR_outpred_d$fit^2),
                                dHAR_pred=sqrt(EU_HAR_outpred_c$fit^2+EU_dHAR_outpred_d$fit^2),
                                naive_pred=sqrt(EU_HAR_outpred_c$fit^2+EU_naive_outpred_d^2))

# Plot the HAR out-of-sample forecast
EU_HAR_outpred_plot_cd <- ggplot() +
  geom_line(data=data.frame(x=EU_daily3_23$date, y=EU_daily3_23$rvol), 
            aes(x = x, y = y, color = "Rvol")) +
  geom_line(data=data.frame(x=EU_HAR_outpred_cd$date, y=EU_HAR_outpred_cd$HAR_pred), 
            aes(x = x, y = y, color = "HAR pred")) +
  geom_line(data=data.frame(x=EU_HAR_outpred_cd$date, y=EU_HAR_outpred_cd$dHAR_pred), 
            aes(x = x, y = y, color = "dHAR pred")) +
  geom_line(data=data.frame(x=EU_HAR_outpred_cd$date, y=EU_HAR_outpred_cd$naive_pred), 
            aes(x = x, y = y, color = "naive pred")) +
  scale_color_manual(name=NULL, values=c('Rvol' = 'darkgrey', "HAR pred" = "blue3", 
                                         "dHAR pred"='steelblue', 'naive pred'='cyan2')) +
  theme_classic() +
  ggtitle("HAR Out-of-sample Forecast (with 95% CI) on diffusive parts of EURUSD from 2023 Mar - Apr") +
  xlab("date") + ylab("BPV")

EU_HAR_outpred_plot_cd
```

## SHAR
### Seperate and explore P/N_RV_c and P/N_RV_d
```{r}
# Obtain daily summary for FX pairs which include the test statistic and corresponding critival values
EU_daily3 <- summary3(EU_flat, lubridate::date, jump_threshold = 0.002)
#AU_daily3 <- summary3(AU_flat, lubridate::date, jump_threshold = 0.005)
#length(EU_daily3$rvol_d[EU_daily3$rvol_d != 0])

acf(EU_daily3$P_rvol_c, lag.max=30, main="EURUSD Daily RVol+ (continous), 2022 May - 2023 Apr")
acf(EU_daily3$P_rvol_d, lag.max=30, main="EURUSD Daily RVol+ (discontinous), 2022 May - 2023 Apr")
acf(EU_daily3$N_rvol_c, lag.max=30, main="EURUSD Daily RVol- (continous), 2022 May - 2023 Apr")
acf(EU_daily3$N_rvol_d, lag.max=30, main="EURUSD Daily RVol- (discontinous), 2022 May - 2023 Apr")
```
### SHAR_RV_c in-sample forecasting for $\sigma^2$
```{r}
EU_SHAR_in_model_c <- comp_shar_pred(23, 256, EU_daily3$rvol_c, EU_daily3$P_rvol_c, EU_daily3$N_rvol_c)$model
EU_SHAR_in_model_c$coefficients
base::summary(EU_SHAR_in_model_c)$adj.r.squared
EU_SHAR_inpred_c <- data.frame(date=EU_daily$date[24:257], pred = predict(EU_SHAR_in_model_c), true = EU_daily3$rvol_c[24:257])

# Plot the HAR in-sample forecast
EU_SHAR_inpred_c_plot <- ggplot(data=EU_SHAR_inpred_c, aes(x=date)) +
  geom_line(aes(y=true, color='True RVol_c')) +
  geom_line(aes(y=pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('True RVol_c' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("RVol") +
  theme_classic()

EU_SHAR_inpred_c_plot
```
### SHAR_RV_c Out-of-sample forecasting for $\sigma^2$
```{r echo=FALSE}
# Obtain the one-day out-of-sample forecast and 0.95 CI of the rolling window
EU_SHAR_outpred_c <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_shar_pred(x-train_length, x, EU_daily3$rvol_c, EU_daily3$P_rvol_c, EU_daily3$N_rvol_c)$pred})), row.names = NULL)
EU_SHAR_outpred_c$date = EU_daily3$date[(EU_roll_start+2) : (EU_roll_end+2)]
EU_SHAR_outpred_c$true = EU_daily3$rvol_c[(EU_roll_start+2) : (EU_roll_end+2)]

mean((EU_SHAR_inpred_c$pred - EU_SHAR_inpred_c$true)^2)
mean((EU_SHAR_outpred_c$true - EU_SHAR_outpred_c$fit)^2)

# Plot the HAR out-of-sample forecast
EU_daily3_23 <- EU_daily3[year(EU_daily3$date)==2023,]
EU_SHAR_outpred_plot_c <- ggplot() +
  geom_line(data=data.frame(x=EU_daily3_23$date, y=EU_daily3_23$rvol_c), 
            aes(x = x, y = y, color = "realized √BPV/µ1")) +
  geom_line(data=data.frame(x=EU_SHAR_outpred_c$date, y=EU_SHAR_outpred_c$fit), 
            aes(x = x, y = y, color = "HAR pred")) +
  geom_ribbon(data=data.frame(x=EU_SHAR_outpred_c$date, lwr=EU_SHAR_outpred_c$lwr, upr = EU_SHAR_outpred_c$upr), aes(x = x, ymin = lwr, ymax=upr), fill = "lightblue", alpha=0.3) +
  scale_color_manual(name=NULL, values=c('realized √BPV/µ1' = 'darkgrey', "HAR pred" = "darkblue")) +
  theme_classic() +
  ggtitle("HAR Out-of-sample Forecast (with 95% CI) on diffusive parts of EURUSD from 2023 Mar - Apr") +
  xlab("date") + ylab("BPV")

EU_SHAR_outpred_plot_c
```
### SHAR, dSHAR and naive in-sample prediction for RV_d
```{r}
# Estimate the model HAR_J
EU_SHAR_in_model_d <- comp_shar_pred(23, 256, EU_daily3$rvol_d, EU_daily3$P_rvol_d, EU_daily3$N_rvol_d)$model
EU_SHAR_in_model_d$coefficients
base::summary(EU_SHAR_in_model_d)$adj.r.squared

# Estimate the model dHAR_J
EU_dSHAR_in_model_d <- comp_dshar_pred(23, 256, EU_daily3$rvol_d, EU_daily3$P_rvol_d, EU_daily3$N_rvol_d)$model
EU_dSHAR_in_model_d$coefficients
base::summary(EU_dSHAR_in_model_d)$adj.r.squared

# Estimate the model Naive_J
EU_naive_in_d <- EU_daily3$rvol_d[23:256]
```
### SHAR RV_c+RV_d in-sample forecast
```{r}
# HAR_BPV + HAR_J
EU_SHAR_inpred_cd <- data.frame(date=EU_daily3$date[24:257], true=EU_daily3$rvol[24:257],
                                SHAR_pred=sqrt(predict(EU_SHAR_in_model_c)^2 +
                                                 predict(EU_SHAR_in_model_d)^2),
                                dSHAR_pred=sqrt(predict(EU_SHAR_in_model_c)^2 +
                                                  predict(EU_dSHAR_in_model_d)^2),
                                naive_pred=sqrt(predict(EU_SHAR_in_model_c)^2 + EU_naive_in_d^2))

# Plot the HAR in-sample forecast
EU_SHAR_inpred_plot_cd <- ggplot(data=EU_SHAR_inpred_cd, aes(x=date)) +
  geom_line(aes(y=true, color='RVol')) +
  geom_line(aes(y=SHAR_pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('RVol' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("BPV+J") +
  theme_classic()

EU_SHAR_inpred_plot_cd

EU_dSHAR_inpred_plot_cd <- ggplot(data=EU_SHAR_inpred_cd, aes(x=date)) +
  geom_line(aes(y=true, color='RVol')) +
  geom_line(aes(y=dSHAR_pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('RVol' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("BPV+J") +
  theme_classic()

EU_dSHAR_inpred_plot_cd

EU_naive_inpred_plot_cd <- ggplot(data=EU_SHAR_inpred_cd, aes(x=date)) +
  geom_line(aes(y=true, color='RVol')) +
  geom_line(aes(y=naive_pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('RVol' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("BPV+J") +
  theme_classic()

EU_naive_inpred_plot_cd
```
### SHAR, dSHAR and naive out-of-sample prediction for RV_d
```{r}
EU_SHAR_outpred_d <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_shar_pred(x-train_length, x, EU_daily3$rvol_d, EU_daily3$P_rvol_d, EU_daily3$N_rvol_d)$pred})), row.names = NULL)

EU_dSHAR_outpred_d <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_dshar_pred(x-train_length, x, EU_daily3$rvol_d, EU_daily3$P_rvol_d, EU_daily3$N_rvol_d)$pred})), row.names = NULL)

EU_naive_outpred_d <- EU_daily3$rvol_d[(EU_roll_start+1): (EU_roll_end+1)]
```

### SHAR RV_c+RV_d out-of-sample forecast
```{r}
# Obtain the one-day out-of-sample forecast of the rolling window
EU_SHAR_outpred_cd <- data.frame(date=EU_daily3$date[(EU_roll_start+2) : (EU_roll_end+2)],
                                 true=EU_daily3$rvol[(EU_roll_start+2) : (EU_roll_end+2)],
                                 SHAR_pred=sqrt(EU_HAR_outpred_c$fit^2+EU_HAR_outpred_d$fit^2),
                                 dSHAR_pred=sqrt(EU_HAR_outpred_c$fit^2+EU_dHAR_outpred_d$fit^2),
                                 naive_pred=sqrt(EU_HAR_outpred_c$fit^2+EU_naive_outpred_d^2))

# Plot the HAR out-of-sample forecast
EU_SHAR_outpred_plot_cd <- ggplot() +
  geom_line(data=data.frame(x=EU_daily3_23$date, y=EU_daily3_23$rvol), 
            aes(x = x, y = y, color = "Rvol")) +
  geom_line(data=data.frame(x=EU_SHAR_outpred_cd$date, y=EU_SHAR_outpred_cd$SHAR_pred), 
            aes(x = x, y = y, color = "HAR pred")) +
  geom_line(data=data.frame(x=EU_SHAR_outpred_cd$date, y=EU_SHAR_outpred_cd$dSHAR_pred), 
            aes(x = x, y = y, color = "dHAR pred")) +
  geom_line(data=data.frame(x=EU_SHAR_outpred_cd$date, y=EU_SHAR_outpred_cd$naive_pred), 
            aes(x = x, y = y, color = "naive pred")) +
  scale_color_manual(name=NULL, values=c('Rvol' = 'darkgrey', "HAR pred" = "blue3", 
                                         "dHAR pred"='steelblue', 'naive pred'='cyan2')) +
  theme_classic() +
  ggtitle("HAR Out-of-sample Forecast (with 95% CI) on diffusive parts of EURUSD from 2023 Mar - Apr") +
  xlab("date") + ylab("BPV")

EU_SHAR_outpred_plot_cd
```