---
title: "Univariate model fitting"
author: "CID: 01714151"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=FALSE, include=FALSE}
y = EU_daily$rvol[24:214]
x_d = EU_daily$rvol[23:213]
comp_mean <- function(index, period, dat){
  idx_end = index
  idx_start = index - period + 1
  return(mean(dat[idx_start:idx_end]))
}
x_w = unlist(lapply(23:213, function(index){comp_mean(index, period=5, dat=EU_daily$rvol)}))
x_m = unlist(lapply(23:213, function(index){comp_mean(index, period=23, dat=EU_daily$rvol)}))
har <- lm(y~x_d + x_w + x_m)
predict(har, data.frame(x_d=EU_daily$rvol[214], x_w=comp_mean(214, 5, EU_daily$rvol), x_m=comp_mean(214, 22, EU_daily$rvol)), interval='confidence', level=0.95)
```

## Define functions for model fitting and forecasting
```{r echo=FALSE}
# Define function that computes RV_{t:t-k} for different periods k
comp_mean <- function(index, period, dat){
  idx_end = index
  idx_start = index - period + 1
  return(mean(dat[idx_start:idx_end]))
}

# Define function that gives the naive one-day forecast
naive_pred <- function(train_start=23, train_end, dat){
  pred = dat[train_start:train_end]
  return(pred)
}

# Define function that computes the one-day forecast of the HAR model
comp_har_pred <- function(train_start=23, train_end, dat){
  # Obtain RV_{t+1} as the training y
  y = dat[(train_start+1):(train_end+1)]
  # Obtain RV_{t}, RV_{t:t-4}, RV_{t:t-22} as the training X
  x_d = dat[train_start:train_end]
  x_w = unlist(lapply(train_start:train_end, function(index){comp_mean(index, period=5, dat=dat)}))
  x_m = unlist(lapply(train_start:train_end, function(index){comp_mean(index, period=23, dat=dat)}))
  # Fit the HAR model with the training model
  har = lm(y~x_d + x_w + x_m)
  # Make prediction on RV_{t+2}
  pred = predict(har, data.frame(x_d=dat[train_end+1], x_w=comp_mean(train_end+1, 5, dat), x_m=comp_mean(train_end+1, 23, dat)), interval='confidence', level=0.95)
  # Return the fitted model and the one-day forecast
  return(list(pred=pred, model=har))
}
# comp_har_pred(23, 213, EU_daily$rvol)

# Define function that computes the one-day forecast of the HAR model of only daily data
comp_dhar_pred <- function(train_start=23, train_end, dat){
  # Obtain RV_{t+1} as the training y
  y = dat[(train_start+1):(train_end+1)]
  # Obtain RV_{t}, RV_{t:t-4}, RV_{t:t-22} as the training X
  x_d = dat[train_start:train_end]
  # Fit the HAR model with the training model
  lm = lm(y~x_d)
  # Make prediction on RV_{t+2}
  pred = predict(lm, data.frame(x_d=dat[train_end+1]), interval='confidence', level=0.95)
  # Return the fitted model and the one-day forecast
  return(list(pred=pred, model=lm))
}

#Define function that computes the one-day forecast of the SHAR model
comp_shar_pred <- function(train_start=23, train_end, dat_d, dat_p, dat_n){
  # Obtain RV_{t+1} as the training y
  y = dat_d[(train_start+1):(train_end+1)]
  # Obtain P(RV_{t}), N(RV_{t}), RV_{t:t-4}, RV_{t:t-22} as the training X
  x_d_p = dat_p[train_start:train_end]
  x_d_n = dat_n[train_start:train_end]
  x_w = unlist(lapply(train_start:train_end, function(index){comp_mean(index, period=5, dat=dat_d)}))
  x_m = unlist(lapply(train_start:train_end, function(index){comp_mean(index, period=23, dat=dat_d)}))
  # Fit the HAR model with the training model
  shar = lm(y~x_d_p + x_d_n + x_w + x_m)
  # Make prediction on RV_{t+2}
  pred = predict(shar, data.frame(x_d_p=dat_p[train_end+1], x_d_n=dat_n[train_end+1] , x_w=comp_mean(train_end+1, 5, dat_d), x_m=comp_mean(train_end+1, 23, dat_d)), interval='confidence', level=0.95)
  return(list(pred=pred, model=shar))
}
# comp_shar_pred(23, 213, EU_daily$rvol, EU_daily$P_rvol, EU_daily$N_rvol)

#Define function that computes the one-day forecast of the SHAR model of only daily data
comp_dshar_pred <- function(train_start=23, train_end, dat_d, dat_p, dat_n){
  # Obtain RV_{t+1} as the training y
  y = dat_d[(train_start+1):(train_end+1)]
  # Obtain P(RV_{t}), N(RV_{t}), RV_{t:t-4}, RV_{t:t-22} as the training X
  x_d_p = dat_p[train_start:train_end]
  x_d_n = dat_n[train_start:train_end]
  # Fit the HAR model with the training model
  shar = lm(y~x_d_p + x_d_n)
  # Make prediction on RV_{t+2}
  pred = predict(shar, data.frame(x_d_p=dat_p[train_end+1], x_d_n=dat_n[train_end+1]), interval='confidence', level=0.95)
  return(list(pred=pred, model=shar))
}
```

## EURUSD
### HAR In-sample forecasting
```{r echo=FALSE}
# Estimate the model on full data sample
EU_HAR_in_model <- comp_har_pred(23, 256, EU_daily$rvol)$model
EU_HAR_in_model$coefficients
EU_HAR_inpred <- data.frame(date=EU_daily$date[24:257], pred=predict(EU_HAR_in_model), true=EU_daily$rvol[24:257])
# Plot the HAR in-sample forecast
EU_HAR_inpred_plot <- ggplot(data=EU_HAR_inpred, aes(x=date)) +
  geom_line(aes(y=true, color='True RVol')) +
  geom_line(aes(y=pred, color='HAR RV pred')) +
  scale_color_manual(name=NULL, values=c('True RVol' = 'darkgrey', "HAR RV pred" = "steelblue")) +
  #ggtitle("HAR in-sample Forecast on EURUSD from 2022 May - 2023 Apr") +
  ylab("RVol") +
  xlab("Date") +
  theme_classic() +
  theme(axis.title = element_text(size= 12.5),
        axis.text = element_text(size= 11.5),
        legend.title = element_text(size= 11),
        legend.text = element_text(size= 11),
        legend.position = c(0.9, 0.85))

EU_HAR_inpred_plot
```
### SHAR in-sample forecasting
```{r echo=FALSE}
# Estimate the model on full data sample
EU_SHAR_in_model <- comp_shar_pred(23, 256, EU_daily$rvol, EU_daily$P_rvol, EU_daily$N_rvol)$model
EU_SHAR_in_model$coefficients
EU_SHAR_inpred <- data.frame(date=EU_daily$date[24:257], pred=predict(EU_SHAR_in_model), true=EU_daily$rvol[24:257])
# Plot the HAR in-sample forecast
EU_SHAR_outpred_plot <- ggplot(data=EU_SHAR_inpred, aes(x=date)) +
  geom_line(aes(y=true, color='True RVol')) +
  geom_line(aes(y=pred, color='SHAR RV pred')) +
  scale_color_manual(name=NULL, values=c('True RVol' = 'darkgrey', "SHAR RV pred" = "darkgreen")) +
  ggtitle("SHAR in-sample Forecast on EURUSD from 2022 May - 2023 Apr") +
  ylab("RVol") +
  xlab("Date") +
  theme_classic() +
  theme(axis.title = element_text(size= 12.5),
        axis.text = element_text(size= 11.5),
        legend.title = element_text(size= 11),
        legend.text = element_text(size= 11),
        legend.position = c(0.9, 0.85))

EU_SHAR_outpred_plot
```
### HAR Out-of-sample forecasting
```{r echo=FALSE}
# Obtain the start and end day of the rolling window
EU_roll_start <- which(EU_daily$date=='2023-02-27')
EU_roll_end <- dim(EU_daily)[1] - 2
# Obtain the number of observations in each training data
train_length <- EU_roll_start - 23
# Obtain the one-day out-of-sample forecast and 0.95 CI of the rolling window
EU_HAR_outpred <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_har_pred(x-train_length, x, EU_daily$rvol)$pred})), row.names = NULL)
EU_HAR_outpred$date = EU_daily$date[(EU_roll_start+2) : (EU_roll_end+2)]
EU_HAR_outpred$true = EU_daily$rvol[(EU_roll_start+2) : (EU_roll_end+2)]

# Plot the HAR out-of-sample forecast
EU_daily_23 <- EU_daily[year(EU_daily$date)==2023,]
EU_HAR_outpred_plot <- ggplot() +
  geom_line(data=data.frame(x=EU_daily_23$date, y=EU_daily_23$rvol), aes(x = x, y = y, color = "True RVol")) +
  geom_line(data=data.frame(x=EU_HAR_outpred$date, y=EU_HAR_outpred$fit), aes(x = x, y = y, color = "HAR pred")) +
  geom_ribbon(data=data.frame(x=EU_HAR_outpred$date, lwr=EU_HAR_outpred$lwr, upr = EU_HAR_outpred$upr), aes(x = x, ymin = lwr, ymax=upr), fill = "lightblue", alpha=0.3) +
  scale_color_manual(name=NULL, values=c('True RVol' = 'darkgrey', "HAR pred" = "darkblue")) +
  theme_classic() +
  ggtitle("HAR Out-of-sample Forecast (with 95% CI) on EURUSD from 2023 Mar - Apr") +
  xlab("date") + ylab("RVol")

EU_HAR_outpred_plot
```
### SHAR out-of-sample forecating
```{r echo=FALSE}
# Obtain the one-day out-of-sample forecast and 0.95 CI of the rolling window
EU_SHAR_outpred <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_shar_pred(x-train_length, x, EU_daily$rvol, EU_daily$P_rvol, EU_daily$N_rvol)$pred})), row.names = NULL)
EU_SHAR_outpred$date = EU_daily$date[(EU_roll_start+2) : (EU_roll_end+2)]
EU_SHAR_outpred$true = EU_daily$rvol[(EU_roll_start+2) : (EU_roll_end+2)]

EU_SHAR_outpred_plot = ggplot() +
  geom_line(data=data.frame(x=EU_daily_23$date, y=EU_daily_23$rvol), aes(x = x, y = y, color = "True RVol")) +
  geom_line(data=data.frame(x=EU_SHAR_outpred$date, y=EU_SHAR_outpred$fit), aes(x = x, y = y, color = "SHAR pred")) +
  geom_ribbon(data=data.frame(x=EU_SHAR_outpred$date, lwr=EU_SHAR_outpred$lwr, upr = EU_SHAR_outpred$upr), aes(x = x, ymin = lwr, ymax=upr), fill = "lightgreen", alpha=0.3) +
  scale_color_manual(name=NULL, values=c('True RVol' = 'darkgrey', "SHAR pred" = "darkgreen")) +
  theme_classic() +
  ggtitle("Semivariance HAR Out-of-sample (with 95% CI) on EURUSD from 2023 Mar-Apr") +
  xlab("date") + ylab("RVol")

EU_SHAR_outpred_plot
```
### Comparison of MSE of HAR and SHAR for both in-sample and out-of-sample forecasting
```{r echo=FALSE}
print(paste("EU_HAR_inpred_MSE:", mean((EU_HAR_inpred$pred - EU_HAR_inpred$true)^2)))
print(paste("EU_SHAR_inpred_MSE:", mean((EU_SHAR_inpred$pred - EU_HAR_inpred$true)^2)))

print(paste("EU_HAR_outpred_MSE:", mean((EU_HAR_outpred$true-EU_HAR_outpred$fit)^2)))
print(paste("EU_SHAR_outpred_MSE:", mean((EU_SHAR_outpred$true-EU_SHAR_outpred$fit)^2)))

print(paste("EU_HAR_R2:", base::summary(EU_HAR_in_model)$adj.r.squared))
print(paste("EU_SHAR_R2:", base::summary(EU_SHAR_in_model)$adj.r.squared))

library(forecast)
dm.test(EU_HAR_inpred$pred - EU_HAR_inpred$true, EU_SHAR_inpred$pred - EU_HAR_inpred$true, alternative='two.sided', power=2)
dm.test(EU_HAR_outpred$true-EU_HAR_outpred$fit, EU_SHAR_outpred$true-EU_SHAR_outpred$fit, alternative='two.sided', power=2)
```

## AURUSD
### HAR In-sample forecasting
```{r echo=FALSE}
# Estimate the model on full data sample
AU_HAR_in_model <- comp_har_pred(23, 256, AU_daily$rvol)$model
AU_HAR_in_model$coefficients
AU_HAR_inpred <- data.frame(date=AU_daily$date[24:257], pred=predict(AU_HAR_in_model), true=AU_daily$rvol[24:257])
# Plot the HAR in-sample forecast
AU_HAR_outpred_plot <- ggplot(data=AU_HAR_inpred, aes(x=date)) +
  geom_line(aes(y=true, color='True RVol')) +
  geom_line(aes(y=pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('True RVol' = 'darkgrey', "in-sample pred" = "orange")) +
  ggtitle("HAR in-sample Forecast on AUDUSD from 2022 May - 2023 Apr") +
  ylab("RVol") +
  theme_classic()

AU_HAR_outpred_plot
```
### SHAR in-sample forecasting
```{r echo=FALSE}
# Estimate the model on full data sample
AU_SHAR_in_model <- comp_shar_pred(23, 256, AU_daily$rvol, AU_daily$P_rvol, AU_daily$N_rvol)$model
AU_SHAR_in_model$coefficients
AU_SHAR_inpred <- data.frame(date=AU_daily$date[24:257], pred=predict(AU_SHAR_in_model), true=AU_daily$rvol[24:257])
# Plot the HAR in-sample forecast
AU_SHAR_outpred_plot <- ggplot(data=AU_SHAR_inpred, aes(x=date)) +
  geom_line(aes(y=true, color='True RVol')) +
  geom_line(aes(y=pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('True RVol' = 'darkgrey', "in-sample pred" = "lightcoral")) +
  ggtitle("SHAR in-sample Forecast on AUDUSD from 2022 May - 2023 Apr") +
  ylab("RVol") +
  theme_classic()

AU_SHAR_outpred_plot
```
### HAR Out-of-sample forecasting
```{r echo=FALSE}
# Obtain the start and end day of the rolling window
AU_roll_start <- which(AU_daily$date=='2023-02-27')
AU_roll_end <- dim(AU_daily)[1] - 2
# Obtain the number of observations in each training data
train_length <- AU_roll_start - 23
# Obtain the one-day out-of-sample forecast and 0.95 CI of the rolling window
AU_HAR_outpred <- data.frame(do.call(rbind, lapply(AU_roll_start:AU_roll_end, function(x){comp_har_pred(x-train_length, x, AU_daily$rvol)$pred})), row.names = NULL)
AU_HAR_outpred$date = AU_daily$date[(AU_roll_start+2) : (AU_roll_end+2)]
AU_HAR_outpred$true = AU_daily$rvol[(AU_roll_start+2) : (AU_roll_end+2)]

# Plot the HAR out-of-sample forecast
AU_daily_23 <- AU_daily[year(AU_daily$date)==2023, ]
AU_HAR_outpred_plot <- ggplot() +
  geom_line(data=data.frame(x=AU_daily_23$date, y=AU_daily_23$rvol), aes(x = x, y = y, color = "True RVol")) +
  geom_line(data=data.frame(x=AU_HAR_outpred$date, y=AU_HAR_outpred$fit), aes(x = x, y = y, color = "HAR pred"), lwd=0.7) +
  geom_ribbon(data=data.frame(x=AU_HAR_outpred$date, lwr=AU_HAR_outpred$lwr, upr = AU_HAR_outpred$upr), aes(x = x, ymin = lwr, ymax=upr), fill = "moccasin", alpha=0.5) +
  scale_color_manual(name=NULL, values=c('True RVol' = 'darkgrey', "HAR pred" = "coral")) +
  theme_classic() +
  ggtitle("HAR Out-of-sample Forecast (with 95% CI) on AUDUSD from 2023 Mar - Apr") +
  xlab("date") + ylab("RVol")

AU_HAR_outpred_plot
```
### SHAR out-of-sample forecating
```{r echo=FALSE}
# Obtain the one-day out-of-sample forecast and 0.95 CI of the rolling window
AU_SHAR_outpred <- data.frame(do.call(rbind, lapply(AU_roll_start:AU_roll_end, function(x){comp_shar_pred(x-train_length, x, AU_daily$rvol, AU_daily$P_rvol, AU_daily$N_rvol)$pred})), row.names = NULL)
AU_SHAR_outpred$date = AU_daily$date[(AU_roll_start+2) : (AU_roll_end+2)]
AU_SHAR_outpred$true = AU_daily$rvol[(AU_roll_start+2) : (AU_roll_end+2)]

AU_SHAR_outpred_plot = ggplot() +
  geom_line(data=data.frame(x=AU_daily_23$date, y=AU_daily_23$rvol), aes(x = x, y = y, color = "True RVol")) +
  geom_line(data=data.frame(x=AU_SHAR_outpred$date, y=AU_SHAR_outpred$fit), aes(x = x, y = y, color = "SHAR pred")) +
  geom_ribbon(data=data.frame(x=AU_SHAR_outpred$date, lwr=AU_SHAR_outpred$lwr, upr = AU_SHAR_outpred$upr), aes(x = x, ymin = lwr, ymax=upr), fill = "lightsalmon", alpha=0.3) +
  scale_color_manual(name=NULL, values=c('True RVol' = 'darkgrey', "SHAR pred" = "lightcoral")) +
  theme_classic() +
  ggtitle("Semivariance HAR Out-of-sample (with 95% CI) on AUDUSD from 2023 Mar - Apr") +
  xlab("date") + ylab("RVol")

AU_SHAR_outpred_plot
```
### Comparison of MSE of HAR and SHAR for both in-sample and out-of-sample forecasting
```{r echo=FALSE}
mean((AU_HAR_outpred$true-AU_HAR_outpred$fit)^2)
mean((AU_SHAR_outpred$true-AU_SHAR_outpred$fit)^2)

mean((AU_HAR_inpred$pred - AU_HAR_inpred$true)^2)
mean((AU_SHAR_inpred$pred - AU_HAR_inpred$true)^2)
```

