---
title: "Univariate jump tests"
author: "CID: 01714151"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Define functions and dataset summary

```{r echo=FALSE}
# Define the constant parameters
mu1 = sqrt(2/pi)
theta = (pi^2/4)+pi-5
sigma = 1/288
# Define the functions which compute the adjusted jump ratio statistic
comp_adj_jump_stat <- function(rvar, rbpv, rqpv){
  J = sigma^(-1/2)/sqrt(theta*pmax(1, rqpv/rbpv^2))*(mu1^(-2)*rbpv/rvar-1)
  return(J)
}

# Construct a summary function to get the realized daily variance, bipower and quadpower
summary2 <- function(dat_flat, time_func){
  dat_summary <- dat_flat %>%
    # Add a column which represents the date/month of the dataset
    mutate(date=time_func(datetime)) %>%
    # Change the index and group by date
    index_by(date) %>%
    group_by(date) %>%
    # get the average price, RVol, P(RVol), N(RVol)
    summarize(currency = first(currency),
              avg_price = mean(price) ,
              rvar = sum(log_return^2),
              P_rvar = sum(log_return[log_return>0]^2),
              N_rvar = sum(log_return[log_return<=0]^2),
              rbpv = sum(bpv),
              rqpv = sigma^(-1)*sum(qpv),
              jump = pmax(0, rvar - mu1^(-2)*rbpv))
  return(dat_summary)
}

# Obtain daily summary for FX pairs which include the test statistic and corresponding critival values
EU_daily2 <- summary2(EU_flat, lubridate::date)
AU_daily2 <- summary2(AU_flat, lubridate::date)

EU_daily2 = EU_daily2 %>%
  mutate(J = comp_adj_jump_stat(rvar, rbpv, rqpv)) %>%
  mutate(ratio_stat = mu1^(-2)*rbpv/rvar) %>%
  mutate(critical_value99 = qnorm(0.01, 1, sqrt(theta*pmax(1, rqpv/rbpv^2))/sigma^(-1/2))) %>%
  mutate(critical_value95 = qnorm(0.05, 1, sqrt(theta*pmax(1, rqpv/rbpv^2))/sigma^(-1/2)))

AU_daily2 = AU_daily2 %>%
  mutate(J = comp_adj_jump_stat(rvar, rbpv, rqpv)) %>%
  mutate(ratio_stat = mu1^(-2)*rbpv/rvar) %>%
  mutate(critical_value99 = qnorm(0.01, 1, sqrt(theta*pmax(1, rqpv/rbpv^2))/sigma^(-1/2))) %>%
  mutate(critical_value95 = qnorm(0.05, 1, sqrt(theta*pmax(1, rqpv/rbpv^2))/sigma^(-1/2)))
```

## EURUSD

### Plot ratio statistic and corresponding adjusted critical values

```{r echo=FALSE}
#EU_jump_stat_plot <- ggplot(EU_daily2) + geom_point(aes(x=date, y=J), shape=3, size=1) + geom_hline(aes(yintercept = qnorm(0.01, 0, 1)), color='steelblue',linetype='dashed') + theme_classic()
#EU_jump_stat_plot

EU_jump_stat_plot <- ggplot(EU_daily2, aes(x=date)) + 
  geom_point(aes(y=ratio_stat, color='ratio statistic'), shape=4, size=1) + 
  geom_line(aes(y=critical_value99, color='99% critical value'), linetype='solid') +
  scale_color_manual(name=NULL, values=c('ratio statistic' = 'black', "99% critical value" = "steelblue"),
                     guide = guide_legend(override.aes = list(
                       linetype = c("ratio statistic" = "solid", "critival value" = "blank"), 
                       shape = c("ratio statistic"= NA, "99% critival value"= 4)))) + 
  #ggtitle("Ratio statistic and corresponding adjusted critical values for EURUSD") +
  xlab("Date") +
  ylab("Ratio statistic value") +
  theme_classic() +
  theme(axis.title = element_text(size= 12.5),
        axis.text = element_text(size= 11.5),
        legend.title = element_text(size= 11),
        legend.text = element_text(size= 11),
        legend.position = c(.14, .15))

EU_jump_stat_plot
```

### Rejection rate

```{r echo=FALSE}
# compute the ratio of rejection at 95% and 99%
dim(EU_daily2[EU_daily2$ratio_stat<EU_daily2$critical_value95, ])[1]/dim(EU_daily2)[1]
dim(EU_daily2[EU_daily2$ratio_stat<EU_daily2$critical_value99, ])[1]/dim(EU_daily2)[1]
# obtain the dates of jumps
EU_daily2[EU_daily2$ratio_stat<EU_daily2$critical_value99, c("date", "rvar", "rbpv", "J")]

EU_daily2$adj.jump = ifelse(EU_daily2$ratio_stat<EU_daily2$critical_value99, EU_daily2$jump, 0)
```

### Case study

2022-07-13 US CPI for June 2022. The US Dollar has set another fresh 19-year-high this morning on the heels of a 9.1% headline CPI print, and is now pulling back. source(<https://www.dailyfx.com/forex/fundamental/daily_briefing/session_briefing/daily_fundamentals/2022/07/13/US-Dollar-Price-Action-Setups-EUR-USD-EURUSD-GBP-USD-GBPUSD-USD-CAD-USDCAD-USD-JPY-USDJPY.html>, <https://www.bls.gov/schedule/2022/home.htm>)

2022-09-13 US CPI for August 2022. EUR plunged for hot US CPI, parity risk. source(<https://www.bls.gov/schedule/news_release/cpi.htm>; <https://www.dailyfx.com/forex/technical/home/analysis/eur-usd/2022/09/13/EURUSD-Short-term-Technical-Outlook-Euro-Post-CPI-Plunge-to-Parity-Trade-Outlook.html>)

2022-10-13 USD CPI for September 2022. source(<https://www.bls.gov/schedule/2022/home.htm>)

2022-11-10 USD CPI for October 2022. USD fall due to CPI well lower than expected. source(<https://www.dailyfx.com/news/forex-fx-gold-stocks-usd-slammed-gold-and-eurusd-jump-after-cpi-20221110.html>)

2022-11-15 EU GDP QoQ and YoY & USD PPI for Octorber 2022. weaker USD, source(<https://www.dailyfx.com/news/forex-eur-usd-nudges-higher-as-euro-area-gdp-hits-estimates-economic-sentiment-beats-20221115.html>)

2022-12-02 US Employment rate for November 2022. source(<https://www.bls.gov/schedule/2022/home.htm>)

```{r echo=FALSE}
# Dates of significant jumps - reject H0 at 0.0001 and large RVar
EU_daily2[(EU_daily2$rvar>1e-4) & (EU_daily2$J < qnorm(0.001, 0, 1)),]

# Take 2022-11-10 as an example - observe y and ratio stats
require(gridExtra)
grid.arrange(
  EU_flat[which((EU_flat$datetime <= "2022-07-14") & (EU_flat$datetime >= "2022-07-12")), ] %>%
  autoplot(log_return) + labs(title = "Log return (y)", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)), 
  ggplot(EU_daily2[which((EU_daily2$date <= "2022-07-14")&(EU_daily2$date >= "2022-07-12")),], aes(x=date)) +
    geom_point(aes(y=ratio_stat), color="black", shape=4, size=2) + 
    geom_line(aes(y=critical_value99), color="steelblue", linetype='solid') +
    labs(title = "Ratio statistic", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)),
  EU_flat[which((EU_flat$datetime <= "2022-09-14") & (EU_flat$datetime >= "2022-09-12")), ] %>%
  autoplot(log_return) + labs(title = "Log return (y)", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)), 
  ggplot(EU_daily2[which((EU_daily2$date <= "2022-09-14")&(EU_daily2$date >= "2022-09-12")),], aes(x=date)) +
    geom_point(aes(y=ratio_stat), color="black", shape=4, size=2) + 
    geom_line(aes(y=critical_value99), color="steelblue", linetype='solid') +
    labs(title = "Ratio statistic", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)),
  EU_flat[which((EU_flat$datetime <= "2022-10-14") & (EU_flat$datetime >= "2022-10-12")), ] %>%
  autoplot(log_return) + labs(title = "Log return (y)", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)), 
  ggplot(EU_daily2[which((EU_daily2$date <= "2022-10-14")&(EU_daily2$date >= "2022-10-12")),], aes(x=date)) +
    geom_point(aes(y=ratio_stat), color="black", shape=4, size=2) + 
    geom_line(aes(y=critical_value99), color="steelblue", linetype='solid') +
    labs(title = "Ratio statistic", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)),
  EU_flat[which((EU_flat$datetime <= "2022-11-11") & (EU_flat$datetime >= "2022-11-09")), ] %>%
    autoplot(log_return) + labs(title = "Log return (y)", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)), 
  ggplot(EU_daily2[which((EU_daily2$date <= "2022-11-11")&(EU_daily2$date >= "2022-11-09")),], aes(x=date)) +
    geom_point(aes(y=ratio_stat), color="black", shape=4, size=2) + 
    geom_line(aes(y=critical_value99), color="steelblue", linetype='solid') +
    labs(title = "Ratio statistic", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)),
  EU_flat[which((EU_flat$datetime <= "2022-11-16") & (EU_flat$datetime >= "2022-11-14")), ] %>%
  autoplot(log_return) + labs(title = "Log return (y)", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)), 
  ggplot(EU_daily2[which((EU_daily2$date <= "2022-11-16")&(EU_daily2$date >= "2022-11-14")),], aes(x=date)) +
    geom_point(aes(y=ratio_stat), color="black", shape=4, size=2) + 
    geom_line(aes(y=critical_value99), color="steelblue", linetype='solid') +
    labs(title = "Ratio statistic", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)),
  EU_flat[which((EU_flat$datetime <= "2022-12-03") & (EU_flat$datetime >= "2022-12-01")), ] %>%
  autoplot(log_return) + labs(title = "Log return (y)", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)), 
  ggplot(EU_daily2[which((EU_daily2$date <= "2022-12-03")&(EU_daily2$date >= "2022-12-01")),], aes(x=date)) +
    geom_point(aes(y=ratio_stat), color="black", shape=4, size=2) + 
    geom_line(aes(y=critical_value99), color="steelblue", linetype='solid') +
    labs(title = "Ratio statistic", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=8.5)),
  ncol=4)
```

### HAR with Bipower vairiation

#### In-sample forecasting

```{r}
# Estimate the model on full data sample
EU_HAR_in_model2 <- comp_har_pred(23, 256, sqrt(EU_daily2$rbpv))$model
EU_HAR_in_model2$coefficients
# Compute adjusted R^2 of the fitted BPV model
base::summary(EU_HAR_in_model2)$adj.r.squared

EU_HAR_inpred2 <- data.frame(date=EU_daily2$date[24:257], pred=predict(EU_HAR_in_model2)*mu1^(-1), true=sqrt(EU_daily2$rbpv[24:257]*mu1^(-2)))
# Plot the HAR in-sample forecast
EU_HAR_inpred_plot2 <- ggplot(data=EU_HAR_inpred2, aes(x=date)) +
  geom_line(aes(y=true, color='realized √BPV/µ')) +
  geom_line(aes(y=pred, color='in-sample pred')) +
  scale_color_manual(name=NULL, values=c('realized √BPV/µ' = 'darkgrey', "in-sample pred" = "steelblue")) +
  ggtitle("HAR in-sample Forecast for diffusive parts on EURUSD from 2022 May - 2023 Apr") +
  ylab("BPV") +
  theme_classic()

EU_HAR_inpred_plot2
```

#### HAR Out-of-sample forecasting
```{r echo=FALSE}
# Obtain the one-day out-of-sample forecast and 0.95 CI of the rolling window
EU_HAR_outpred2 <- data.frame(do.call(rbind, lapply(EU_roll_start:EU_roll_end, function(x){comp_har_pred(x-train_length, x, sqrt(EU_daily2$rbpv))$pred*mu1^(-1)})), row.names = NULL)
EU_HAR_outpred2$date = EU_daily2$date[(EU_roll_start+2) : (EU_roll_end+2)]
EU_HAR_outpred2$true_bpv = sqrt(EU_daily2$rbpv[(EU_roll_start+2) : (EU_roll_end+2)]*mu1^(-2))

mean((EU_HAR_inpred2$pred - EU_HAR_inpred2$true)^2)
mean((EU_HAR_outpred2$true_bpv-EU_HAR_outpred2$fit)^2)

# Plot the HAR out-of-sample forecast
EU_daily2_23 <- EU_daily2[year(EU_daily2$date)==2023,]
EU_HAR_outpred_plot2 <- ggplot() +
  geom_line(data=data.frame(x=EU_daily2_23$date, y=sqrt(EU_daily2_23$rbpv*mu1^(-2))), 
            aes(x = x, y = y, color = "realized √BPV/µ")) +
  geom_line(data=data.frame(x=EU_HAR_outpred2$date, y=EU_HAR_outpred2$fit), 
            aes(x = x, y = y, color = "HAR pred")) +
  geom_ribbon(data=data.frame(x=EU_HAR_outpred2$date, lwr=EU_HAR_outpred2$lwr, upr = EU_HAR_outpred2$upr), aes(x = x, ymin = lwr, ymax=upr), fill = "lightblue", alpha=0.3) +
  scale_color_manual(name=NULL, values=c('realized √BPV/µ' = 'darkgrey', "HAR pred" = "darkblue")) +
  theme_classic() +
  ggtitle("HAR Out-of-sample Forecast (with 95% CI) on diffusive parts of EURUSD from 2023 Mar - Apr") +
  xlab("date") + ylab("BPV")

EU_HAR_outpred_plot2
```





## AUDUSD

### Plot ratio statistic and corresponding adjusted critical values

```{r echo=FALSE}
AU_jump_stat_plot <- ggplot(AU_daily2, aes(x=date)) + 
  geom_point(aes(y=ratio_stat, color='ratio statistic'), shape=3, size=1) + 
  geom_line(aes(y=critical_value99, color='99% critical value'), linetype='solid') +
  scale_color_manual(name=NULL, values=c('ratio statistic' = 'black', "99% critical value" = "orange"),
                     guide = guide_legend(override.aes = list(
                       linetype = c("ratio statistic" = "solid", "critival value" = "blank"), 
                       shape = c("ratio statistic"= NA, "99% critival value"= 3)))) + 
  ggtitle("Ratio statistic and corresponding adjusted critical values for AUDUSD") +
  xlab("Date") +
  ylab("ratio statistic magnitude") +
  theme_classic() +
  theme(legend.position = c(.15, .15))

AU_jump_stat_plot
```

```{r echo=FALSE}
# require(gridExtra)
# plot two graphs together
# grid.arrange(EU_jump_stat_plot, AU_jump_stat_plot, ncol=2)
```

### Rejection rate

```{r echo=FALSE}
dim(AU_daily2[AU_daily2$ratio_stat<AU_daily2$critical_value95, ])[1]/dim(AU_daily2)[1]
dim(AU_daily2[AU_daily2$ratio_stat<AU_daily2$critical_value99, ])[1]/dim(AU_daily2)[1]
```


```{r}
grid.arrange(
  EU_flat[which((EU_flat$datetime <= "2022-07-14") & (EU_flat$datetime >= "2022-07-12")), ] %>%
    autoplot(log_return) + labs(title = "Log return (y)", y=NULL, x=NULL) + 
    theme_classic() + theme(axis.text.x = element_text(size=10)),
  ggplot(EU_daily2[which((EU_daily2$date <= "2022-07-14")&(EU_daily2$date >= "2022-07-12")),], aes(x=date)) + 
    geom_point(aes(y=ratio_stat, color="ratio statistic"), shape=4, size=2) + 
    geom_line(aes(y=critical_value99, color='99% critical value'), linetype='solid') +
  scale_color_manual(name=NULL, values=c('ratio statistic' = 'black', "99% critical value" = "steelblue"),
                     guide = guide_legend(override.aes = list(
                       linetype = c("ratio statistic" = "solid", "critival value" = "blank"), 
                       shape = c("ratio statistic"= NA, "99% critival value"= 4)))) +
    labs(title = "Ratio statistic", y=NULL, x=NULL) + 
    theme_classic() + 
    theme(axis.text.x = element_text(size=10), 
          legend.position = c(.8, .8)), ncol=2)
```

```{r eval=FALSE, include=FALSE}
ggplot(data.frame(date=EU_daily2$date, 
                  jump=EU_daily2$jump, tranformed_jump=log(EU_daily2$jump+1e-6)), 
       aes(x = date)) +
  geom_line(aes(y=jump, color='jump')) +
  geom_line(aes(y= tranformed_jump*(1e-4), color='tranformed_jump')) +
  scale_y_continuous(name = "jump",
                     sec.axis = sec_axis(trans= ~ ./1e-4, name="tranformed_jump")) +
  scale_color_manual(name=NULL, values=c('jump' = 'steelblue', "tranformed_jump" = "orange")) +
  #ggtitle("Daily Average Rate of Currency Pairs from 2022 May - 2023 Apr") +
  xlab("Date") +
  theme_classic() +
  theme(axis.title = element_text(size= 12.5),
        axis.text = element_text(size= 11.5),
        legend.title = element_text(size= 11),
        legend.text = element_text(size= 11),
        legend.position = c(0.85, 0.15))
```
```{r}
grid.arrange(
ggplot(data.frame(date=EU_daily2$date, 
                  jump=EU_daily2$jump, tranformed_jump=log(EU_daily2$jump+1e-6)), 
       aes(x = date)) +
  geom_line(aes(y=jump, color='J = max(0, RV-BV)')) +
  scale_color_manual(name=NULL, values=c('J = max(0, RV-BV)' = 'steelblue')) +
  #ggtitle("Daily Average Rate of Currency Pairs from 2022 May - 2023 Apr") +
  xlab("Date") + ylab("jump") +
  theme_classic() +
  theme(axis.title = element_text(size= 12.5),
        axis.text = element_text(size= 11.5),
        legend.title = element_text(size= 11),
        legend.text = element_text(size= 11),
        legend.position = c(0.85, 0.85)),
ggplot(data.frame(date=EU_daily2$date, 
                  jump=EU_daily2$jump, tranformed_jump=log(EU_daily2$jump+1e-6)), 
       aes(x = date)) +
  geom_line(aes(y= tranformed_jump, color='log(J+1e-6)')) +
  scale_color_manual(name=NULL, values=c('log(J+1e-6)' = 'darkgreen')) +
  #ggtitle("Daily Average Rate of Currency Pairs from 2022 May - 2023 Apr") +
  xlab("Date") + ylab("transformed jump") +
  theme_classic() +
  theme(axis.title = element_text(size= 12.5),
        axis.text = element_text(size= 11.5),
        legend.title = element_text(size= 11),
        legend.text = element_text(size= 11),
        legend.position = c(0.85, 0.85)), 
ggplot(data.frame(date=EU_daily2$date, 
                  jump=EU_daily2$jump, tranformed_jump=log(EU_daily2$jump+1e-10)), 
       aes(x = date)) +
  geom_line(aes(y= tranformed_jump, color='log(J+1e-10)')) +
  scale_color_manual(name=NULL, values=c('log(J+1e-10)' = 'coral')) +
  #ggtitle("Daily Average Rate of Currency Pairs from 2022 May - 2023 Apr") +
  xlab("Date") + ylab("transformed jump") +
  theme_classic() +
  theme(axis.title = element_text(size= 12.5),
        axis.text = element_text(size= 11.5),
        legend.title = element_text(size= 11),
        legend.text = element_text(size= 11),
        legend.position = c(0.85, 0.85)), nrow=3)
```

